---
title: "hmw2_problems"
author: "shuai yuan"
date: "2024-09-29"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(janitor)


```




## Problem 1
Import subway E&T data and clean data

list.files()

## Load and Clean Data

```{r }
# List available files in the working directory
list.files()

# Load the dataset
subway_data <- read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

# Select relevant columns and clean data
cleaned_subway_data <- subway_data %>%
  select(Line, `Station Name`, `Station Latitude`, `Station Longitude`, 
         Route1, Route2, Route3, Route4, Route5, 
         Entry, Vending, `Entrance Type`, ADA) %>%
  mutate(Entry = ifelse(Entry == "YES", TRUE, FALSE)) %>%  # Convert 'Entry' to logical values
  distinct()  # Remove duplicates

# Check the dimensions of the cleaned dataset
dim(cleaned_subway_data)


```

## Q1: How many distinct stations are there (by station name and line)?

```{r}
num_distinct_stations <- cleaned_subway_data %>%
  distinct(Line, `Station Name`) %>%
  nrow()

cat("Number of distinct stations:", num_distinct_stations, "\n")

```
## Q2: How many stations are ADA compliant?
```{r }
num_ada_compliant_stations <- cleaned_subway_data %>%
  filter(ADA == TRUE) %>%
  distinct(Line, `Station Name`) %>%
  nrow()

cat("Number of ADA compliant stations:", num_ada_compliant_stations, "\n")

```
## Q3: What proportion of entrances/exits without vending allow entry?
```{r }
proportion_no_vending_allow_entry <- cleaned_subway_data %>%
  filter(Vending == "NO") %>%
  summarise(proportion = mean(Entry)) %>%
  pull(proportion)

cat("Proportion of no-vending entrances that allow entry:", proportion_no_vending_allow_entry, "\n")

```


## Reformat data: combine all route columns into a single list
```{r }

cleaned_subway_data <- cleaned_subway_data %>%
  unite("Routes", Route1:Route5, sep = ",", na.rm = TRUE, remove = FALSE) %>%
  separate_rows(Routes, sep = ",")

```
## Q4 How many distinct stations serve the A train?
```{r }
num_a_train_stations <- cleaned_subway_data %>%
  filter(Routes == "A") %>%
  distinct(Line, `Station Name`) %>%
  nrow()

cat("Number of stations serving the A train:", num_a_train_stations, "\n")
```
# Q5: Of the stations that serve the A train, how many are ADA compliant?

```{r }
num_ada_compliant_a_train_stations <- cleaned_subway_data %>%
  filter(Routes == "A", ADA == TRUE) %>%
  distinct(Line, `Station Name`) %>%
  nrow()

cat("Number of ADA compliant stations serving the A train:", num_ada_compliant_a_train_stations, "\n")

```
## show results


```{r pressure, echo=FALSE}
cat("Number of distinct stations:", num_distinct_stations, "\n")
cat("Number of ADA compliant stations:", num_ada_compliant_stations, "\n")
cat("Proportion of no-vending entrances that allow entry:", proportion_no_vending_allow_entry, "\n")
cat("Number of stations serving the A train:", num_a_train_stations, "\n")
cat("Number of ADA compliant stations serving the A train:", num_ada_compliant_a_train_stations, "\n")
```
In Problem 1, we successfully imported, cleaned, and analyzed the NYC Subway Entrance & Exit dataset. By filtering and transforming the data, we identified that there are a significant number of distinct subway stations across various lines. We also examined the accessibility of these stations and found the number of ADA-compliant stations. Additionally, we calculated the proportion of no-vending entrances that still allow entry, providing insights into the availability of entry options for passengers. 

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Problem 2

## /Users/mac/Downloads/202309 Trash Wheel Collection Data.xlsx
Import data of Mr trash wheel
```{r}
mr_trash_wheel_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N586",
             na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    sports_balls = as.numeric(round(sports_balls, digits = 0)),
    trash_wheel = "mr_trash_wheel",
    year = as.double(year)
  )

```
## show head 

```{r}
head(mr_trash_wheel_df, 5)
```

## Import and Clean Data for Professor Trash Wheel


```{r}

prof_trash_wheel_df <- read_excel("./202409 Trash Wheel Collection Data.xlsx",
                                  sheet = "Professor Trash Wheel",
                                  range = "A2:M108",
                                  na = c("NA", "", ".")) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(trash_wheel = "prof_trash_wheel")

head(prof_trash_wheel_df, 5)


```


## Import and Clean Data for Gwynnda Trash Wheel


```{r}

gwynnda_trash_wheel_df <- read_excel("./202409 Trash Wheel Collection Data.xlsx",
                                     sheet = "Gwynnda Trash Wheel",
                                     range = "A2:L157",
                                     na = c("NA", "", ".")) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(trash_wheel = "gwynnda_trash_wheel")

head(gwynnda_trash_wheel_df, 5)

```
## Combine All Trash Wheel Data


```{r}

trash_wheel_tidy <- bind_rows(mr_trash_wheel_df, prof_trash_wheel_df, gwynnda_trash_wheel_df) %>%
  janitor::clean_names() %>%
  relocate(trash_wheel)


```


##Calculate Total Weight of Professor Trash Wheel and Cigarette Butts Collected by Gwynnda in June 2022


```{r}

total_weight = sum(trash_wheel_tidy$weight_tons[trash_wheel_tidy$trash_wheel == "prof_trash_wheel"],
                   na.rm = TRUE)

cigarette_butt = sum(trash_wheel_tidy$cigarette_butts[
  trash_wheel_tidy$trash_wheel == "gwynnda_trash_wheel" &
  trash_wheel_tidy$month == "June" &
  trash_wheel_tidy$year == 2022],
  na.rm = TRUE)


```

##Results

The total weight of trash collected by Professor Trash Wheel is r **total_weight tons**, and the total number of cigarette butts collected by Gwynnda Trash Wheel in June of 2022 is r **cigarette_butt counts**. 

### Problem 2 detailss:
- **`clean_names()`** from the `janitor` package is used to clean the column names.
- **`read_excel()`** specifies the sheet, range, and missing value handling for each Trash Wheel dataset.
- We filter out rows where `dumpster` is `NA`, and the `mutate()` function is used to handle numeric transformations (e.g., rounding sports balls, converting year to numeric).
- The datasets are combined with `bind_rows()` and then cleaned again for consistency.


#Problem 3


## Import and clean 
```{r}

bakers_df = 
  read_csv("./gbb_datasets/bakers.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  separate(
    baker_name, into = c("baker", "last_name"), sep = " "
  )
head(bakers_df, 5)

# Import and clean the bakes data
bakes_df = 
  read_csv("./gbb_datasets/bakes.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  mutate(
    baker = replace(baker, baker == '"Jo"', "Jo")  # Fix any potential formatting issues with names
  )
head(bakes_df, 5)

# Import and clean the results data
results_df = 
  read_csv("./gbb_datasets/results.csv", skip = 2, na = c("NA", "", ".")) |>
  janitor::clean_names()
head(results_df, 5)

```

## merging the datasets

```{r}

# Merge the datasets
baker_show_df = left_join(results_df, bakers_df, by = c("baker", "series")) |>
  left_join(bakes_df, by = c("baker", "series", "episode")) |>
  relocate(baker, last_name, baker_age, baker_occupation, hometown, signature_bake, show_stopper) |>
  arrange(baker) |>
  mutate(result = str_to_lower(result)) |>
  filter(!is.na(result))  # Ensure that only rows with result are kept
head(baker_show_df, 5)

# Export the merged and cleaned dataset
write_csv(baker_show_df, "./gbb_datasets/baker_show.csv")


```



## Star Baker and Winners from Seasons 5 to 10

```{r}

# Filter to get winners and star bakers from Seasons 5 to 10
baker_winner = 
  baker_show_df |>
  filter(series >= 5 & series <= 10) |>
  filter(result == "winner" | result == "star baker") |>
  relocate(series, episode, result) |>
  arrange(series, episode)

# Display the filtered results
head(baker_winner, 5)


```

## 5. Import Viewership Data


```{r}
# Import and clean the viewership data
viewers_df = 
  read_csv("./gbb_datasets/viewers.csv", na = c("NA", "", ".")) |>
  janitor::clean_names()

# Show the first 10 rows of the viewership data
head(viewers_df, 10)



```
## Average Viewership for Seasons 1 and 5


```{r}

# Calculate average viewership for Season 1 and Season 5
session_1 = round(mean(viewers_df$series_1, na.rm = TRUE), 2)
session_5 = round(mean(viewers_df$series_5, na.rm = TRUE), 2)

# Display the results
session_1
session_5


```

##conclusion

The average viewership in Season 1 was r session_1 million, while the average viewership in Season 5 was r session_5 million. This indicates a significant increase in popularity over the seasons.





